---
title: "Homework 2"
format: html
editor: source
---

<hr style="border: 3px solid black;" />

## Task 1: Basic Vector Practice

#### 1. Pre and Post treatment vectors

```{r}
pre_treatment <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)

post_treatment <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
```

#### 2. Renamed each

```{r}
names(pre_treatment) <- paste("Subject", 1:length(pre_treatment), sep = "_")
names(post_treatment) <- paste("Subject", 1:length(post_treatment), sep = "_")

pre_treatment
post_treatment
```

#### 3. Difference from pre -> post treatment

```{r}
delta <- pre_treatment - post_treatment
```

#### 4. Average change - positive value (positive outcome, blood pressure decreased on average by this amount)

```{r}
mean(delta)
```

#### 5. Which subjects had a decreased blood pressure?

```{r}
decrease <- names(which(delta > 0))
decrease
```

#### 6. Which subjects had an decrease in blood pressure ( positive change, but as a subset vector)?

```{r}
decrease_vector <- delta[delta > 0]
decrease_vector
```

#### 7. What is the average decrease of blood pressure of those that had the positive change?

```{r}
mean(decrease_vector)
```

<hr style="border: 3px solid black;" />

## Task 2: Basic Data Frame Practice

#### 1. Create dataframe
```{r}
data <- data.frame(
  patient = names(pre_treatment),
  pre_bp = pre_treatment,
  post_bp = post_treatment,
  diff_bp = delta,
  row.names=NULL
)
head(data, 1)
```

#### 2. Show where diff_bp is negative
```{r}
data[data$diff_bp < 0,]
```

#### 3. Add new column where post_bp is less than 120 (T/F)
```{r}
data$"post_bp < 120" = data$post_bp < 120
data
```

#### 4. Pretty print view using provided function
```{r}
knitr::kable(data, align='c', caption = "Patient Treatment BP effect")
```

<hr style="border: 3px solid black;" />

## Task 3: List Practice

#### 1. Create similar dataframe to Task 2
```{r}
df <- data.frame(
  patient = paste("Subject", 1:10, sep = "_"),
  pre_bp = c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130),
  post_bp = c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124),
  row.names = NULL
)
df$diff_bp <- df$pre_bp - df$post_bp
df$"post_bp < 20" <- df$post_bp < 120
head(data, 1)
head(df, 1)
```

#### 2. Store the 2 data frames into a list.
```{r}
my_list <- list(treatment=data, placebo=df)
my_list
```

#### 3. Access the first item 3 different ways.
```{r}
head(my_list[1], 1) # keeps as list (will show entire item)
head(my_list[[1]], 1) # grabs the value at the list position, giving a dataframe - head prints out the first item of df
head(my_list$treatment, 1) # same thing, different way of getting the dataframe
```

#### 4. From list created in question 2, access pre_bp column of placebo list data. Maintaining as df column, as opposed to $pre_bp vector.
```{r}
my_list$placebo["pre_bp"]
```

