---
title: "Homework 1"
format: html
editor: source
---

<hr style="border: 3px solid black;" />

### What do you think being a data scientist is about?

A data scientist is someone who focuses on the process of integrating computers with statistics and business needs. They have a specialization of computing software to be able to gain insights from data, many times in an automated fashion for use with software.

### What differences/similarities do you see between data scientists and statisticians?

I view it as starting on the same path and diverging at some point. The root is in statistics, but a statistician will continue with more of the rigor of the field directly, whereas a data scientist will begin to branch off more into the software/development process and application side of things.

Similarities:

-   Foundational Knowledge
-   Tool Use

Differences:

-   Specialized Knowledge
-   Development Process

### How do you view yourself in relation to these two areas?

I am much closer to a data scientist. I have exposure to statistics, but as portions of CSC courses. I have had many courses on directly attributed data science topics and I work in software.

<hr style="border: 3px solid black;" />

#### Setting y variable using a column of the iris dataset and density function

```{r}
y <- density(iris$Petal.Width)
```

<br>

#### Checking the class, type, and structure of y

```{r}
class(y)
typeof(y)
str(y)
```

<br>

#### Plotting y, with code hidden with echo=FALSE

```{r}
#| echo=FALSE
plot(y)
```
